Research Cases

### 2025-12-16 Code/R/Research/cases_research.Rmd
Add polynomial modulation cases

# compare_mssa_rec <- function() {
#   noise <- rnorm(N * P, 0, sd)
#   x <- signal + noise
#   
#   rec <- list(
#     ssa = reconstruct(ssa(x, L = L_mssa, kind="mssa"), groups = groups_mssa),
#     hossa = tens_mssa_reconstruct(
#       x,
#       L_homssa,
#       groups = groups_homssa,
#       groups3 = groups3_homssa,
#       decomp = "hooi",
#       status = FALSE
#     )
#   )
#   
#   sapply(rec, function(m)
#     Reduce("+", m))
# }
```

@@

# true is nd-array, pred is (n+1)d-array with n+1-mode slices as predictions
rmse_ts_nd <- function(true, pred) {
  sqrt(mean(apply(pred, length(dim(pred)), \(p) mse(true, p))))
}

## true is vector (ts), pred is matrix with columns as predictions of the ts
# rmse_ts <- function(true, pred) {
#   # sqrt(mean(apply(pred, 2, \(p) mean((true - p)^2))))
#   sqrt(mean(apply(pred, 2, \(p) mse(true, p))))
# }

## true is matrix (mts) with columns as individual ts, pred is 3d-array with 3-mode slices as predictions of the mts
# rmse_mts <- function(true, pred) {
#   sqrt(mean(apply(pred, 3, \(p) mse(true, p))))  
# }
```

@@

# params_grid_list: list with the parameter names as names and vectors of the parameters values as values
calc_errors_by_params <- function(params_grid_list, R, comp_func, err_func, signal, seed = 5) {
  params_grid <- expand.grid(params_grid_list)
  results <- params_grid
  results$error <- numeric(nrow(params_grid))
  q <- progressor(nrow(params_grid) * R)
  for (i in seq(nrow(params_grid))) {
    # print(paste0(names(params_grid), " = ", params_grid[i,], collapse = ", "))
    list2env(params_grid[i,], envir = parent.frame())
    results[i,]$error <- calc_errors(R, comp_func, err_func, signal, q, seed)
    # q()
  }
  return(results)
}
```

```{r}
# If parameters are matrices, then a multivariate series will be constructed.
# In this parameters matrices columns correspond to summnands and rows to channels
# This means, if you want to increase number of ts in a sum, you grow columns,
# and if you want to add a channel to an mv ts, you grow rows
# The results is either a 1d ts, or a matrix, where each column is a 1d ts -- channel.
construct_periodics <- function(N,
                                ampl,
                                rate,
                                freq,
                                phase,
                                complex = TRUE,
                                from = 0,
                                by = 1) {
  if (any(c(length(ampl), length(rate), length(freq)) != length(phase))) {
    stop("Amplitudes, rates, frequencies and phases all must have the same length")
  }
  if (all(is.matrix(ampl), is.matrix(rate), is.matrix(freq), is.matrix(phase))) {
    if (any(c(dim(ampl), dim(rate), dim(freq)) != dim(phase))) {
      stop("Amplitudes, rates, frequencies and phases all must have the same dimensions")
    }
  } else if (any(is.matrix(ampl), is.matrix(rate), is.matrix(freq), is.matrix(phase))) {
    stop("All parameters should be either matrices, or vectors at the same time")
  }

  if (complex) {
    period_f <- exp
    unit <- 1i
  } else {
    period_f <- cos
    unit <- 1
  }
  N_ones <- rep(1, N)
  ts_arr <- (N_ones %o% ampl) * (N_ones %o% exp(rate)) *
    period_f(N_ones %o% phase + 2 * pi * unit * seq(
      from = from,
      length.out = N,
      by = by
    ) %o% freq)

@@

calc_rank <- function(ampl, rate, freq, phase, complex = TRUE, tol = 1e-6) {
  if (any(c(length(ampl), length(rate), length(freq)) != length(phase)))
    stop("Amplitudes, rates, frequencies and phases all must have the same length")
  fr_unique <- unique(matrix(c(freq, rate), ncol = 2))
  (2 - complex) *
    (sum(ampl != 0) + nrow(fr_unique) - length(ampl)) - sum(abs(fr_unique[,1] - 0.5) < tol)
}

@@

## Parameters for P from 2 to 20 with one periodic
```{r}
# P <- 2
# P_case <- "_c2"
# P <- 4
# P_case <- "_c4"
# P <- 8
# P_case <- "_c8"
# P <- 12
# P_case <- "_c12"
# P <- 16
# P_case <- "_c16"
P <- 20
P_case <- "_c20"

@@

@@

```{r}
full_case <- paste0(n_case, P_case, rates_case, freqs_case, A_case, phases_case, complex_case)
max.r <- calc_rank(A, rates.true, freqs.true, phases.true, complex.signal)

stopifnot(N > max.r * 2 + 1)
signal <- construct_periodics(N, A, rates.true, freqs.true, phases.true, complex.signal)
```

@@

if (n_case == "") {
  Ls_mssa <- seq(from = max.r + 1,
                to = N,
                by = 2)
  Ls_homssa = seq(from = max.r + 1,
                to = (N + 1) %/% 2,
                by = 2)
} else {
  Ls_mssa <- seq(from = max.r + 1,
                to = N,
                by = 10)
  Ls_homssa = seq(from = max.r + 1,
                to = (N + 1) %/% 2,
                by = 10)
}

>>>>

  Reduce("+", rec)
}
```

@@

# true is nd-array, pred is (n+1)d-array with n+1-mode slices as predictions
rmse_ts_nd <- function(true, pred) {
  sqrt(mean(apply(pred, length(dim(pred)), \(p) mse(true, p))))
}
```

@@

# params_grid_list: list with the parameter names as names and vectors of the parameters values as values
calc_errors_by_params <- function(params_grid_list, R, comp_func, err_func, signal, seed = 5) {
  params_grid <- expand.grid(params_grid_list)
  results <- params_grid
  results$error <- numeric(nrow(params_grid))
  q <- progressor(nrow(params_grid) * R)
  # q <- progressor(nrow(params_grid))
  for (i in seq(nrow(params_grid))) {
    # print(paste0(names(params_grid), " = ", params_grid[i,], collapse = ", "))
    list2env(params_grid[i,], envir = parent.frame())
    results[i,]$error <- calc_errors(R, comp_func, err_func, signal, q, seed)
    # q()
  }
  return(results)
}
```

```{r}
polynom_eval <- function(coefs, x) {
  powers <- outer(x, seq_along(coefs) - 1, FUN = "^")
  result <- powers %*% coefs

  return(as.vector(result))
}

# If parameters are matrices, then a multivariate series will be constructed.
# In this parameters matrices columns correspond to summands and rows to channels
# This means, if you want to increase number of ts in a sum, you grow columns,
# and if you want to add a channel to an mv ts, you grow rows
# The result is either a 1d ts, or a matrix, where each column is a 1d ts -- channel.
construct_periodics <- function(N,
                                ampl,
                                rate,
                                freq,
                                phase,
                                complex = TRUE,
                                poly_ampl_coefs = NULL,
                                from = 0,
                                by = 1) {
  if (any(c(length(ampl), length(rate), length(freq)) != length(phase))) {
    stop("Amplitudes, rates, frequencies and phases all must have the same length")
  }
  if (all(is.matrix(ampl), is.matrix(rate), is.matrix(freq), is.matrix(phase))) {
    if (any(c(dim(ampl), dim(rate), dim(freq)) != dim(phase))) {
      stop("Amplitudes, rates, frequencies and phases all must have the same dimensions")
    }
    if (is.null(poly_ampl_coefs)) {
      poly_ampl_coefs <- 1 %o% rep(1, nrow(ampl)) %o% rep(1, ncol(ampl))
    } else {
      stopifnot(
                is.array(poly_ampl_coefs) && 
                length(dim(poly_ampl_coefs)) - 1 == length(dim(ampl)) &&
                all(dim(poly_ampl_coefs)[-1] == dim(ampl))
               )
    }
  } else {
    if (any(is.matrix(ampl), is.matrix(rate), is.matrix(freq), is.matrix(phase))) {
      stop("All parameters should be either matrices, or vectors at the same time")
    }

    if (is.null(poly_ampl_coefs)) {
      poly_ampl_coefs <- 1 %o% rep(1, length(ampl))
    } else {
      stopifnot(
                is.matrix(poly_ampl_coefs) && 
                length(dim(poly_ampl_coefs) == 2) &&
                dim(poly_ampl_coefs)[2] == length(ampl)
               )
    }
  }

  if (complex) {
    period_f <- exp
    unit <- 1i
  } else {
    period_f <- cos
    unit <- 1
  }
  N_ones <- rep(1, N)
  ts_range <- seq(from = from, length.out = N, by = by)
  poly_ampl <- apply(poly_ampl_coefs, 2:length(dim(poly_ampl_coefs)), polynom_eval, x=ts_range)
  ts_arr <- poly_ampl * (N_ones %o% ampl) * (N_ones %o% exp(rate)) *
    period_f(N_ones %o% phase + 2 * pi * unit * ts_range %o% freq)

@@

calc_rank <- function(ampl, rate, freq, phase, complex = TRUE, poly_ampl_coefs = NULL, tol = 1e-6) {
  if (any(c(length(ampl), length(rate), length(freq)) != length(phase)))
    stop("Amplitudes, rates, frequencies and phases all must have the same length")
  if (is.null(poly_ampl_coefs))
    pac_rank <- rep(1, length(ampl))
  else
    pac_rank <- apply(poly_ampl_coefs, 2:length(dim(poly_ampl_coefs)), function(v) max(which(v != 0)))
  fr_df <- data.frame(freq = freq, rate = rate, pac = pac_rank)
  fr_unique <- fr_df |> group_by(freq, rate) |> summarise(pac = max(pac))
  (2 - complex) *
    (sum(ampl != 0) + sum(fr_unique$pac) - length(ampl)) - sum(abs(fr_unique[,1] - 0.5) < tol)
}

@@

## Parameters for P from 2 to 20 with one periodic
```{r}
set.seed(5)

P <- 2
P_case <- "_c2"
# P <- 4
# P_case <- "_c4"
# P <- 8
# P_case <- "_c8"
# P <- 12
# P_case <- "_c12"
# P <- 16
# P_case <- "_c16"
# P <- 20
# P_case <- "_c20"

@@

# poly.true <- 1 %o% rep(1, P) %o% 1
# poly_case <- ""
poly.true <- c(1, 1) * matrix(runif(2 * P, 0.5, 2), nrow = 2) %o% 1
poly_case <- "p1"

@@

```{r}
full_case <- paste0(n_case, P_case, rates_case, freqs_case, A_case, poly_case, phases_case, complex_case)
max.r <- calc_rank(A, rates.true, freqs.true, phases.true, complex.signal, poly.true)

stopifnot(N > max.r * 2 + 1)
signal <- construct_periodics(N, A, rates.true, freqs.true, phases.true, complex.signal, poly.true)
```

@@

if (n_case == "") {
  Ls_mssa <- seq(from = max.r + 1,
                to = N - max.r - 1,
                by = 2)
  Ls_homssa = seq(from = max.r + 1,
                to = (N + 1) %/% 2,
                by = 2)
} else {
  Ls_mssa <- seq(from = max.r + 1,
                to = N - max.r - 1,
                by = 10)
  Ls_homssa = seq(from = max.r + 1,
                to = (N + 1) %/% 2,
                by = 10)
}
